---
title       : Showing surface uncertainty using smooth animations
subtitle    : 
author      : Charles R. Hogg III
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [mathjax, bootstrap]     # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}

---

# Stuff I want to communicate!

The **point** is we need a way to view a probability distribution over surfaces.
(It's one thing to compute the distribution; we also need to _visualize_ it.)

**Early on**: a *critical* idea:
  - The input to a multivariate Gaussian is a bunch of independent 1-D Gaussians
    - Plot a 
  - So we can focus on getting a good 1D Gaussian, which is easier to visualize!

---

```{r setup, include=FALSE}
require(knitr)
require(rgl)
knit_hooks$set(webgl = function(before, options, envir) {
  library(rgl)
  ## after a chunk has been evaluated
  if (before || rgl.cur() == 0) return()  # no active device
  name <- sprintf('rgl_%s.html', options$label)
  par3d(windowRect = 100 + options$dpi * c(0, 0, options$fig.width, options$fig.height))
  Sys.sleep(.05) # need time to respond to window size change
  writeLines(c('%WebGL%', '<script>webGLStart();</script>'), tpl <- tempfile())
  writeWebGL(dir = dirname(name), filename = name, template = tpl)
  sprintf('<iframe src="%s" width="%s" height="%s"></iframe>', name,
    options$dpi * options$fig.width, options$dpi * options$fig.height)
})
```

---

```{r test3d_setup, echo=FALSE, include=FALSE}
require(gppois)
set.seed(2)

# Generate synthetic datapoints and plotting grid
low.noise <- 1e-6
actual.noise <- 0.2 ^ 2
X.data <- HexagonalGrid(n=3, unit=3)
m.true <- Model("true")
data.0 <- data.frame(X=X.data, Y=0)
d.0 <- Dataset(id='throwaway', data=data.0,
    X.names=c('X.1', 'X.2'), column='Y')
c.true <- CovarianceSE(ell=5, sigma.f=0.2)
m.true$AddCovariance(c.true)
m.true$SetNoiseBounds(low.noise)
K <- gppois:::KTotal.Model(m.true, d.0)
L <- t(chol(K))
Y.true <- L %*% rnorm(d.0$n)
Y.noisy <- Y.true + actual.noise * rnorm(d.0$n)
data <- Dataset(id='random', data=data.frame(X=X.data, Y=Y.noisy),
  X.names=c('X.1', 'X.2'), column='Y')
X.plot <- HexagonalGrid(n=30, unit=0.3)
N.plot <- nrow(X.plot)
data$Plot2D()

# Create gppois models to train on datapoints
m.wide <- Model(id="wide")
m.wide$AddCovariance(CovarianceSE(ell.bounds=c(5, 10), sigma.f.bounds=c(0.01, 1000)))
m.wide$SetNoiseBounds(actual.noise)
m.jumpy <- Model(id="jumpy")
m.jumpy$AddCovariance(CovarianceSE(ell.bounds=c(0.2, 1), sigma.f.bounds=c(0.01, 1000)))
m.jumpy$SetNoiseBounds(actual.noise)
m.wide$Train(d=data)
m.jumpy$Train(d=data)

# Make the plots
p.wide <- m.wide$PosteriorInterval(d=data, X.out=X.plot)
p.jumpy <- m.jumpy$PosteriorInterval(d=data, X.out=X.plot)
PlotSurface(X=X.plot, Y=p.wide$mean, color='blue', new.window=FALSE)
PlotSurface(X=X.plot, Y=p.jumpy$mean, color='red', new.window=FALSE)


# Matrices for plotting draws
L.wide <- m.wide$L(d=data, X.out=X.plot)
L.jumpy <- m.jumpy$L(d=data, X.out=X.plot)
```

---

## Introduction

We want to estimate **surfaces**.
  - (Plot some surface)
So, we have a probability distribution whose elements are these guys; this one here happens to be the mean.

Now, it's one thing to _compute_ a probability distribution; quite another to
_visualize_ it.



---

### First try: pointwise uncertainty

One common approach is to give some kind of credible interval at each point.
These usually won't vary much from point to point, so they come out looking
pretty smooth.  (Here's a 1D example for visual clarity.)

The problem is, you have thrown away any information about how the points
relate to each other. This information is critically important; for instance,
in this case, actual members of the distribution look more like this, or this;
the credible interval approach can mislead.

Not to mention, in 2D, it can be kind of hard to see the credible intervals.

---

```{r pointwise, webgl=TRUE, echo=FALSE}
data$Plot2D()
L.wide <- m.wide$L(d=data, X.out=X.plot)
PlotSurface(X=X.plot, Y=(L.wide %*% rnorm(N.plot)), color='blue', new.window=FALSE)
L.jumpy <- m.jumpy$L(d=data, X.out=X.plot)
PlotSurface(X=X.plot, Y=(L.jumpy %*% rnorm(N.plot)), color='red', new.window=FALSE)
```

---

### Second try: plotting draws from the distribution

HERE

---

### Existing Solution

ESG 1997.
  - Pro:
    - Elegant, obviously-correct interpolation
    - Computationally fast
  - Con:
    - Chained to keyframes
      - Abrupt velocity changes

Good segueway into the advantages of mine...

Advantages of ESG which mine retains:
  - Simple, elegant, exact
Disadvantages of ESG which mine corrects:
  - perfectly smooth motion
  - no longer chained to keyframes!

_One line of `R` code!_\
...with `n` semicolons\
(Okay, it's `n+1` lines of `R` code.)

Getting the statistics **right** -- that's **critical**!
  - Show animated Q-Q plots to drive home this point

---

## Continuous animation

  1. Naive interpolation: _wrong statistics_
  2. ESG: elegant and correct
  3. Making it smooth

---

### First pass: connect the dots

(show draws from normal distribution; function of time)

(show dots connected)

(show tons of timetraces; highlight one)

So what's wrong with this?

(Plot animated Q-Q plots)

Basis functions

---

# Other ideas

Atomic motion might be a good way to show the advantages of smoothness

---

## Hey, is ESG only independent at the keyframes?

Might be good to point out.

---

## Conclusions

- Part of my goal is to make ESG's excellent work more widely known
- Animated uncertainty visualization belongs in the toolbox of every statistical researcher.
- If that uncertainty has a Gaussian distribution, the method I've presented is an excellent way to achieve it.
